Comparison of Global Solution Methods to a Zero Lower Bound Model

### 2019-04-25 Paper/Paper.tex
Presentation done; just need to add featured bullet pts

In this paper, we consider the effect of making the policy functions conditional on whether the ZLB binds. To isolate the effect of splitting the policy functions, we use policy function iteration on a fixed point with linear interpolation following \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}. This solution method also provides the motivation for a single policy function, and the \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution provides the motivation for a regime-indexed policy function. Henceforth, the solution method based on the \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}} solution algorithm will be denoted ART and the solution algorithm that estimates smoother functions based on \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} will be denoted GHLS. It is important to note that we do not consider the effect of a Smolyak basis with Chebyshev polynomials, so this is a modified \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution method. The key point in this paper is to explore how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model.

@@

To approximate regime-indexed consumption policy functions $c_t^{ZLB}$ and $c_t^{normal}$, we provide initial guesses based on the log-linear solution. We refine the policy function guesses as follows: For each node in the state space, we solve for the time $t$ variables with either $c_t^{ZLB}$  or $c_t^{normal}$ according to $i_t$, the interest rate at time $t$. With the time $t$ variables, we linearly interpolate both policy functions. Next, we calculate the interest rate on each node in the state space, construct an aggregate consumption policy function based on whether the economy is at the ZLB on each node, and calculate time $t+1$ variables accordingly. We then update $c_t^{ZLB}$ with an interest rate of 1, and $c_t^{normal}$ with $i_t$.

@@

\noindent
\textbf{Monetary Policy} The central bank sets the gross nominal interest rate, $i$, according to
\begin{gather}
    \label{eq:15}
    i_t=\max\{1,i_t^n\},\\
      \label{eq:16}
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{g}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}), 0 \leq \rho_i < 1, \varepsilon_i \sim \mathds{N}(0,1),
  \end{gather}
where $y^{gdp}$ is real GDP (i.e., output, $y$, minus the resources lost due to adjustment costs, $adj^p$), $i^n$ is the gross notional interest rate, $\bar i$ and $\bar{\pi}$ are the target values of the inflation and nominal interest rates, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. A more negative net notional rate indicates that the central bank is more constrained.\\

@@

\begin{table}[H]
  \hypertarget{Table 1}
\centering
\captionsetup{justification=centering}
  \small
    \setlength{\tabcolsep}{7.5pt}      
  \begin{tabular}{l c c l c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.005$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.0035$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
  \caption{Parameter values}
  \label{Tab:1}
\end{table}

\subsection{Parameter Values} \hyperlink{Table 1}{Table 1} shows the model parameters. Parameters are from \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}, and are chosen to be characteristic of U.S. data. The subjective discount factor, $\beta$ is set to $0.9949$, the time average of the values implied by the steady-state consumption Euler equation and the federal funds rate. The Frisch labor supply elasticity $1/\eta$ is set to 3 to match the macro estimate in \hyperlink{Peterman}{\color{black}{Peterman (2016)}}. The price elasticity of substitution $\theta$ is set to 6, matching GHLS and corresponding to to a 20\% average markup. Steady-state labor hours $\bar{n}$ is set to 1/3 of available time. The steady-state risk premium, growth rate, and inflation rate are set to the time averages of per capital real DGP growth, the percent GDP implicit price deflator, and the Baa corporate bond yield relative to the yield on the 10-Year Treasury. The capital share of income $\alpha$ is equal to the \hyperlink{Fernald}{\color{black}{Fernald (2012)}} utilization-adjusted quarterly-TFP estimates of the capital share of income from 1988-2017Q4. The other parameters are set to values consistent with posterior estimates from similar models in the literature. For the model without capital, all parameters are equivalent to those above except we set $\sigma_s=0.006$ and $\sigma_i=0.002$ allow the GHLS algorithm to converge.

@@

where $\mathds{I}_t(d)$ is defined by
\begin{gather}
    \label{eq:19}
  \mathds{I}_t(d) = 
\begin{cases}
     1 &\text{ if } i_t > 1\\
     0 &\text{ otherwise.}
\end{cases}
\end{gather}
The variable $i_t=\max\{1,i_t^n\}$ represents the value of the notional rate derived from evaluating the functions $\textbf{pf}_{t,j}(d)$ for $j \in \{1,2\}$ and using equation (\ref{eq:16}). (For each variable, $j=1$ denotes the function associated with the regime with a positive nominal rate and $j=2$ denotes the function associated with the ZLB regime.) %In calculating the fixed point, we use $i=1$ when the model is in the ZLB regime. 
The functions $\textbf{pf}_{t,j}$ satisfy the residual functions $R_{t,l,j}$ for and $l \in \{1,2,3,4\}$ which correspond to the household FOC bond, FOC capital, FOC investment, and the price Phillips curve, respectively, and $j \in \{1,2\}$.
\begin{gather}
\label{eq:20}
  R_{t,1,1} = 1 - s_ti_t\beta E_t[(\lambda_t/\lambda_{t+1})(1/(\bar{\pi}\pi_{t+1}^{gap}))],\\
\label{eq:21}
  R_{t,1,2} = 1 - s_t\beta E_t[(\lambda_t/\lambda_{t+1})(1/(\bar{\pi}\pi_{t+1}^{gap}))],\\
\label{eq:22}
R_{t,2,j} = q_t - \beta E_t[(\lambda_t/\lambda_{t+1})(r^k_{t+1}+(1-\delta)q_{t+1})],\\
 \label{eq:23} 
  \begin{split}
    R_{t,3,j} &= 1 - q_t[1-\nu(x^g_t-1)^2/2 - \nu(x_t^g-1)x_t^g] \\&- \nu\beta\bar{g}E_t[q_{t+1}(\lambda_t/\lambda_{t+1})(x^g_{t+1})^2(x^g_{t+1}-1)], 
  \end{split}
  \\
   \label{eq:24}
  \begin{split}
    R_{t,4,j} &= \varphi(\pi_t^{gap}-1)\pi_t^{gap} - (1-\theta) - \theta mc_t \\&- \beta\varphi E_t[(\lambda_t/\lambda_{t+1})(\pi_{t+1}^{gap}-1)\pi_{t+1}^{gap}(y_{t+1}/y_t)].
    \end{split}
  \end{gather}

@@

\begin{table}[H]
  \centering
\captionsetup{justification=centering}
  \hypertarget{Table 2}
  \small
    \setlength{\tabcolsep}{6pt}      
  \begin{tabular}{l c c c c c c}
    \hline
    & \multicolumn{3}{c}{Model without capital} & \multicolumn{3}{c}{Model with capital} \\
    & Total nodes & Iterations & Total Time & Total Nodes & Iterations & Total time\\
    \hline
  ART & 2401  & 76 & 21.6s & 78125 & 65 & 0h 15m 37.1s\\    
  GHLS & 2401  & 62 & 18.5s & 78125 & 1170 &  5h 32m 48.6s\\
  \hline
\end{tabular}
  \caption{Solution times}
\end{table}

@@

For the model without capital, the GHLS policy functions do not experience as pronounced a nonlinearity.

@@

\hyperlink{Figure 2}{Figure 2} shows the cross-section of the consumption policy functions with the interest rate and
risk premium. The policy functions are all fairly smooth and only have a slight nonlinearity at the
ZLB. The GHLS policy functions are smoother than the ART policy function with respect to both the average percent error from a linear plane and the RMSE.

\begin{table}[H]
  \centering
\captionsetup{justification=centering}
  \hypertarget{Table 3}
  \small
    \setlength{\tabcolsep}{10pt}      
  \begin{tabular}{l c c c c}
    \hline
    & \multicolumn{2}{c}{Model without capital} & \multicolumn{2}{c}{Model with capital} \\
    & Mean \% Error & RMSE & Mean \% Error & RMSE\\
    \hline
    ART policy & 0.64407\% &0.0027327 $c$  units & 0.201367\% & 0.0029924 $l$ units \\    
  GHLS policy & 0.62694\% & 0.0026704 $c$ units & 0.732262\% & 0.010799 $l$ units \\
  \hline
\end{tabular}
  \caption{Smoothness measures for labor policy functions ($c$ for model with capital and $n$ for model with capital). GHLS combined policy functions are reported.}
\end{table}

@@

\begin{figure}[H]
\hypertarget{Figure 3}  
    \centering
\captionsetup{justification=centering}
    \includegraphics[width=16cm,scale=.8]{eeerrorsCAP.pdf}%{pfs3D_CAPITAL.eps}
  \caption{Euler equation errors for model with capital}    
  \end{figure}

@@

The Euler equation errors are very similar between solution algorithm, though the ART
Euler equation errors are slightly more negative, and thus correspond to a smaller household error.
The mean errors are similar, though the GHLS maximum errors are larger. The shapes of the errors are comparable.

For the model without capital, GHLS performs slightly better.

@@

\hyperlink{Figure 4}{Figure 4} shows the distribution of the absolute value of the Euler equation errors in base 10
logarithms for the consumption Euler equation and the price Phillips curve. The mean and maximum Euler equations are comparable between algorithms, although GHLS has slightly smaller errors.

>>>>

In this paper, we consider the effect of making the policy functions conditional on whether the ZLB binds. To isolate the effect of splitting the policy functions, we use policy function iteration on a fixed point with linear interpolation following \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}. This solution method directly computes the policy functions, while the \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution method indirectly computes the policy functions through regime-indexed policy functions. Henceforth, the solution method based on the \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}} solution algorithm will be denoted ART and the solution algorithm that estimates smoother functions based on \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} will be denoted GHLS. It is important to note that we do not consider the effect of a Smolyak basis with Chebyshev polynomials, so this is a modified \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution method. The key point in this paper is to explore how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model.

@@

To approximate regime-indexed consumption policy functions $c_t^{ZLB}$ and $c_t^{normal}$, we provide initial guesses based on the log-linear solution. We refine the policy function guesses as follows: For each node in the state space, we solve for the time $t$ variables with either $c_t^{ZLB}$  or $c_t^{normal}$ according to $i_t$, the interest rate at time $t$. With the time $t$ variables, we linearly interpolate both policy functions. Next, we calculate the interest rate on each node in the state space, construct an aggregate consumption policy function based on whether the economy is at the ZLB on each node, and calculate time $t+1$ variables accordingly. We then update $c_t^{ZLB}$ with the interest rate constrained at the ZLB, and $c_t^{normal}$ with a positive interest rate.

@@

\noindent
\textbf{Monetary Policy} The central bank sets the gross nominal interest rate, $i$, according to
\begin{gather}
    \label{eq:15}
    i_t=\max\{1,i_t^n\},\\
      \label{eq:16}
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{gdp}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}), 0 \leq \rho_i < 1, \varepsilon_i \sim \mathds{N}(0,1),
  \end{gather}
where $y^{gdp}$ is real GDP (i.e., output, $y$, minus the resources lost due to adjustment costs, $adj^p$), $i^n$ is the gross notional interest rate, $\bar i$ is the target values of the nominal interest rate, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. A more negative net notional rate indicates that the central bank is more constrained.\\

@@

\begin{table}[H]
  \hypertarget{Table 1}
\centering
\captionsetup{justification=centering}
  \small
    \setlength{\tabcolsep}{7.5pt}      
  \begin{tabular}{l c c l c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.0085$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.002$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
  \caption{Parameter values}
  \label{Tab:1}
\end{table}

\subsection{Parameter Values} \hyperlink{Table 1}{Table 1} shows the model parameters. Parameters are from \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}, and are chosen to be characteristic of U.S. data. The subjective discount factor, $\beta$ is set to $0.9949$, the time average of the values implied by the steady-state consumption Euler equation and the federal funds rate. The Frisch labor supply elasticity $1/\eta$ is set to 3 to match the macro estimate in \hyperlink{Peterman}{\color{black}{Peterman (2016)}}. The price elasticity of substitution $\theta$ is set to 6, matching GHLS and corresponding to to a 20\% average markup. Steady-state labor hours $\bar{n}$ is set to 1/3 of available time. The steady-state risk premium, growth rate, and inflation rate are set to the time averages of per capital real DGP growth, the percent GDP implicit price deflator, and the Baa corporate bond yield relative to the yield on the 10-Year Treasury. The capital share of income $\alpha$ is equal to the \hyperlink{Fernald}{\color{black}{Fernald (2012)}} utilization-adjusted quarterly-TFP estimates of the capital share of income from 1988-2017Q4. The other parameters are set to values consistent with posterior estimates from similar models in the literature. For the model without capital, all parameters are equivalent to those above except we set $\sigma_s=0.006$ to allow the GHLS algorithm to converge.

@@

where $\mathds{I}_t(d)$ is defined by
\begin{gather}
    \label{eq:19}
  \mathds{I}_t(d) = 
\begin{cases}
     1 &\text{ if } i_t > 1\\
     0 &\text{ otherwise.}
\end{cases}
\end{gather}
The variable $i_t=\max\{1,i_t^n\}$ represents the value of the notional rate derived from evaluating the functions $\textbf{pf}_{t,j}(d)$ for $j \in \{1,2\}$ and using equation (\ref{eq:16}). (For each variable, $j=1$ denotes the function associated with the regime with a positive nominal rate and $j=2$ denotes the function associated with the ZLB regime.) %In calculating the fixed point, we use $i=1$ when the model is in the ZLB regime. 
The functions $\textbf{pf}_{t,j}$ satisfy the residual functions $R_{t,l,j}$ for $l \in \{1,2,3,4\}$ which correspond to the household FOC bond, FOC capital, FOC investment, and the price Phillips curve, respectively, and $j \in \{1,2\}$.
\begin{gather}
\label{eq:20}
  R_{t,1,1} = 1 - s_ti_t\beta E_t[(\lambda_t/\lambda_{t+1})(1/(\bar{\pi}\pi_{t+1}^{gap}))],\\
\label{eq:21}
  R_{t,1,2} = 1 - s_t\beta E_t[(\lambda_t/\lambda_{t+1})(1/(\bar{\pi}\pi_{t+1}^{gap}))],\\
\label{eq:22}
R_{t,2,j} = q_t - \beta E_t[(\lambda_t/\lambda_{t+1})(r^k_{t+1}+(1-\delta)q_{t+1})],\\
 \label{eq:23} 
  \begin{split}
    R_{t,3,j} &= 1 - q_t[1-\nu(x^g_t-1)^2/2 - \nu(x_t^g-1)x_t^g] \\&- \nu\beta\bar{g}E_t[q_{t+1}(\lambda_t/\lambda_{t+1})(x^g_{t+1})^2(x^g_{t+1}-1)], 
  \end{split}
  \\
   \label{eq:24}
  \begin{split}
    R_{t,4,j} &= \varphi(\pi_t^{gap}-1)\pi_t^{gap} - (1-\theta) - \theta mc_t \\&- \beta\varphi E_t[(\lambda_t/\lambda_{t+1})(\pi_{t+1}^{gap}-1)\pi_{t+1}^{gap}(y_{t+1}/y_t)].
    \end{split}
  \end{gather}

@@

\begin{table}[H]
  \centering
\captionsetup{justification=centering}
  \hypertarget{Table 2}
  \small
    \setlength{\tabcolsep}{6pt}      
  \begin{tabular}{l c c c c c c}
    \hline
    & \multicolumn{3}{c}{Model without capital} & \multicolumn{3}{c}{Model with capital} \\
    & Total nodes & Iterations & Total Time & Total Nodes & Iterations & Total time\\
    \hline
  ART & 2401  & 76 & 23.01s (avg. 30 runs) & 78125 & 56 & 0h 9m 16.1s\\    
  GHLS & 2401  & 62 & 19.34s (avg. 30 runs) & 78125 & 1231 &  3h 28m 28.8s\\
  \hline
\end{tabular}
  \caption{Solution times}
\end{table}

@@

@@

\hyperlink{Figure 2}{Figure 2} shows the cross-section of the consumption policy functions with the interest rate and
risk premium. The policy functions are all fairly smooth and only have a slight nonlinearity at the
ZLB. The GHLS policy functions do not experience a profound nonlinearity as they do in the model with capital. The GHLS policy functions are smoother than the ART policy function with respect to both the average percent error from a linear plane and the RMSE.

\begin{table}[H]
  \centering
\captionsetup{justification=centering}
  \hypertarget{Table 3}
  \small
    \setlength{\tabcolsep}{10pt}      
  \begin{tabular}{l c c c c}
    \hline
    & \multicolumn{2}{c}{Model without capital} & \multicolumn{2}{c}{Model with capital} \\
    & Mean \% Error & RMSE & Mean \% Error & RMSE\\
    \hline
    ART policy & 0.64407\% &0.0027327 $c$  units & 0.271154\% & 0.0039806 $l$ units \\    
  GHLS policy & 0.62694\% & 0.0026704 $c$ units & 0.578098\% & 0.0080785 $l$ units \\
  \hline
\end{tabular}
  \caption{Smoothness measures for labor policy functions ($c$ for model with capital and $n$ for model with capital). GHLS combined policy functions are reported.}
\end{table}

@@

\begin{figure}[H]
\hypertarget{Figure 3}  
    \centering
\captionsetup{justification=centering}
    \includegraphics[width=16cm,scale=.8]{eeerrorsCAP.pdf}%{pfs3D_CAPITAL.pdf}
  \caption{Euler equation errors for model with capital}    
  \end{figure}

@@

The Euler equation errors are comparable between solution algorithms. The household FOC bond and the price Phillips curve are slightly more negative thought the Euler equation errors are slightly more negative in general, and thus correspond to a smaller household error. The shapes of the errors are similar. %though the ART
%Euler equation errors are slightly more negative, and thus correspond to a smaller household error.
%The mean errors are similar, though the GHLS maximum errors are larger.

@@

\hyperlink{Figure 4}{Figure 4} shows the distribution of the absolute value of the Euler equation errors in base 10
logarithms for the consumption Euler equation and the price Phillips curve. The mean and maximum Euler equations are comparable between algorithms, although GHLS has slightly smaller errors. Unlike the model with capital, GHLS performs better.
